Revisiting the hydrodynamic modulation of short surface waves by longer waves

### 2025-03-12 paper.tex
Final edits for revision

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that affects remote sensing, the
interpretation of \textit{in situ} wave measurements, and numerical wave
forecasting.
In this paper, we revisit the wave crest and action conservation laws and derive
higher-order steady solutions for the change of short-wave wavenumber, action,
and gravitational acceleration due to the presence of longer waves.
We validate the analytical solutions with numerical simulations of the full wave
crest and action conservation equations.
The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
steepness modulation significantly deviate from the linear analytical solutions
of \citet{longuet1960changes}, and are similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
The short-wave steepness modulation is attributed to be 5/8 due to the
wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
Examining the homogeneity and stationarity requirements for the conservation of
wave action reveals that stationarity is a stronger requirement and is
generally not satisfied for very steep long waves.
We further examine the results of \citet{peureux2021unsteady} who found through
numerical simulations that the short-wave modulation grows unsteadily with
each long-wave passage.
We show that this unsteady growth only occurs for homogeneous initial
conditions as a special case and does not generally occur, for example in more
realistic long-wave groups.
The proposed steady solutions are a good approximation of the fully
nonlinear numerical solutions in long-wave steepness up to $\approx$0.2.
Except for a subset of initial conditions, the solutions to the non-linearised
crest-action conservation equations are mostly steady in the reference frame of
the long waves.
\end{abstract}

@@

\begin{equation}
\label{eq:lhs1960wavenumber}
\widetilde{k} = k (1 + \varepsilon_L \cos{\psi})
\end{equation}
where $\widetilde{k}$ and $k$ are the modulated and unmodulated short-wave wavenumbers,
respectively, and $\varepsilon_L = a_L k_L$ is the steepness of the long wave with
wavenumber $k_L$ and amplitude $a_L$, and $\psi$ is the long-wave phase.
This process is illustrated in Fig. \ref{fig:hydrodynamic_modulation_diagram}.
Hydrodynamic modulation is a key process for the development and
interpretation of remote sensing of ocean surface waves and currents
\citep{keller1975microwave,alpers1978two,hara1994hydrodynamic}, and is distinguished
from aerodynamic modulation \citep{donelan1987effect,belcher1999wave,chen2000effects}.
Experimental studies of the general short-wave modulation problem included both
the laboratory \citep[e.g.,][]{keller1975microwave,donelan2010modulation} and field
\citep[e.g.,][]{hara2003observation,plant1977studies} measurements.
\citet{phillips1981dispersion} and \citet{longuet1987propagation} revisited the
hydrodynamic modulation problem by considering nonlinear long waves and
variations in the effective gravity of short waves.
Their results yielded significantly stronger modulation than previously predicted
by \citet{longuet1960changes}.
\citet{henyey1988energy} derived an analytical solution for the modulation of
short waves using Hamiltonian mechanics and reported similar modulation magnitudes
to that of \citet{longuet1987propagation}.
\citet{zhang1990evolution} considered weakly nonlinear short waves
using a nonlinear Schr\" odinger equation and found stronger wavenumber
modulation but weaker amplitude modulation than those predicted by
\citet{longuet1987propagation}.
Aside from the early linear solutions by \citet{longuet1960changes}, the variety
of analytical and numerical frameworks (Eulerian crest and action conservation,
Hamiltonian, Schr\" odinger) that tackle the same fundamental
physics and that relatively closely reproduce the steady solutions for short-wave
modulation suggests that the solutions are robust.
Nevertheless, alternative analytical and numerical approaches to the modulation
problem remain of interest, especially if higher degrees of nonlinearity can be
considered using simpler approaches.

@@

All modulation solutions mentioned thus far are steady in the reference frame
of the long wave---short waves become shorter and higher (and thus, steeper)
preferentially on the long-wave crests.
Conversely, they elongate and flatten preferentially in the long-wave troughs.
Recently, \citet{peureux2021unsteady} asked whether the steady (stationary) solutions
of short-wave modulation are appropriate by simulating the full wave conservation
equations.
They found that the short waves grow unsteadily due to the propagation of
longer waves, suggesting that the steady-state solutions may only be valid for
a few long-wave periods before the short waves begin to break due to excessive
steepness.
However, the unsteadiness of their solutions occurs in a specific scenario in
which a long-wave train suddenly appears to modulate a uniform field of short
waves.
The numerical simulations stabilise if the short-wave field is initialised
from existing steady solutions, like those by \citet{longuet1960changes}.
Nevertheless, their results do put into question whether the short-wave
modulation predicted by theory is generally steady over many long-wave periods
or not.
Looking at the conservation equations alone, it is not obvious that it is.

\begin{table}
\begin{center}
\def~{\hphantom{0}}
\begin{tabular}{cccccc}
Reference   & Analysis   & Gravity & Nonlinear  & Nonlinear   & Solution \\
            & framework & varies  & long waves & short waves &          \\
\hline
\citet{longuet1960changes} & Velocity potential  & No  & No  & No  & Analytical \\
\hline
\citet{phillips1981dispersion} & Crest and action & Yes & Yes & No  & Numerical \\
                               & conservation     &     &     &     &           \\
\hline
\citet{longuet1987propagation} & Crest and action & Yes & Yes & No  & Numerical \\
                               & conservation     &     &     &     &           \\
\hline
\citet{henyey1988energy} & Hamiltonian & Yes & Yes & No  & Analytical \\
\hline
\citet{zhang1990evolution} & Schr\" odinger & Yes & Yes & Yes & Numerical \\
\hline
\citet{peureux2021unsteady} & Crest and action & Yes & No  & No  & Numerical \\
                            & conservation     &     &     &     &           \\
\hline
This paper  & Crest and action & Yes & Yes & No  & Analytical \\
            & conservation     &     &     &     & and numerical \\
\hline
\end{tabular}
\caption{
  Non-exhaustive summary of prior and present (this paper) approaches to
  calculating the modulation of short waves by long waves.
}
\label{table:modulation_literature_summary}
\end{center}
\end{table}

In this paper, we revisit this problem from the perspective of wave action
conservation but consider the long-wave slope to be significant enough to
require evaluating the short-waves at the long-wave surface rather than at the
mean water level.
This allows us to derive alternative, steady, nonlinear solutions for the
modulation of short waves, within the limits of small $\varepsilon_L$.
The analytical solutions yield modulations that are stronger than the original
solution by \citet{longuet1960changes}, but similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution} in
long-wave steepness up to $\approx 0.2$.
To validate the steadiness of the analytical solutions, we perform numerical
simulations of the full wave crest and action conservation laws, and compare
the results to the analytical solutions presented here, as well as the prior
numerical results.
The numerical simulations are used to determine the validity of the approximate
steady solution and to investigate the unsteadiness of the short-wave modulation
found by \citet{peureux2021unsteady}.
The main prior studies on hydrodynamic modulation, and their assumptions and
approaches to the solution, are summarised in Table
\ref{table:modulation_literature_summary}.
We first present the governing equations in \S\ref{section:governing_equations},
and then the analytical solutions for the modulation of short waves by long waves
in \S\ref{section:analytical_solutions}.
Then, we describe the numerical simulations and their results in
\S\ref{section:numerical_solutions}.
Finally, we discuss the results and their implications in
\S\ref{section:discussion} and conclude the paper in \S\ref{section:conclusions}.

@@

\begin{equation}
\label{eq:dispersion}
\omega = \sqrt{gk} + k U
\end{equation}
where $\omega$ is the angular frequency in a fixed reference frame, $g$ is the
gravitational acceleration, $k$ is the wavenumber, and $U$ is the mean advective
current in the direction of the wave propagation.
A current in the direction of the waves increases their absolute frequency
without changing their wavenumber.
An important limitation here is that $U$ must be slowly varying on the scales of
the wave period \citep{bretherton1968wavetrains}.
From the perspective of short waves riding on long waves, the advective current
is the horizontal near-surface orbital velocity of the long wave.
Strictly, a velocity profile must be considered when determining the
advective velocity for a wave of any given wavenumber \citep{stewart1974hf},
however for brevity here we will assume a surface velocity to be the advective
one for short waves.
The evolution of the short-wave wavenumber is described by the conservation of
wave crests \citep{phillips1981dispersion}:

@@

\begin{equation}
\label{eq:wave_action}
\dfrac{\partial N}{\partial t}
+ \dfrac{\partial}{\partial x} \left[\left(C_g + U\right)N\right]
= 0
\end{equation}
where $N$ is the action of short waves defined as the ratio of their
energy density to their intrinsic frequency, $E/\sigma$,
and $C_g = 1/2\sqrt{g/k}$ is their group speed in deep water.
The wave action expressed as $E/\sigma$ is commonly taken as
conservative in wave modulation theory
\citep{phillips1981dispersion,longuet1987propagation}, however as
\citet{dysthe1988orbiting} show, it is conservative only for gently sloped waves.
We describe this requirement on gentle wave slope in more detail further below.
Although the absence of wave growth and dissipation is not generally realistic in
the open ocean, it is a useful approximation here to isolate the effects
of hydrodynamic modulation solely due to the motion of the long waves.

@@

\begin{equation}
\label{eq:wave_action2}
\dfrac{\partial N}{\partial t}
+ \left(C_g + U\right) \dfrac{\partial N}{\partial x}
+ N \dfrac{\partial U}{\partial x}
+ N \dfrac{\partial C_g}{\partial x}
= 0
\end{equation}
This equation is similar to (\ref{eq:wavenumber}), except for the last term
which represents the inhomogeneity of the group speed.
As \citet{peureux2021unsteady} explained, the presence of this term causes
unsteady growth of wave action in infinite long-wave trains.
(\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are the governing equations
that we approximate and solve analytically in \S\ref{section:analytical_solutions},
and numerically in their full form in \S\ref{section:numerical_solutions}.
It is worth noting that this system of equations is semi-coupled, meaning that
it is possible to solve (\ref{eq:wavenumber}) on its own, but solving
(\ref{eq:wave_action2}) requires also the solution of (\ref{eq:wavenumber}).
In other words, the wave kinematics (the wavenumber distribution) are not
concerned with the wave action.
In contrast, the wave action is governed, among others, by the
group velocity and thus depends on the wavenumber.

@@

We now describe the analytical solutions for the short-wave wavenumber,
effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
and intrinsic phase speed in presence of long waves.
The wavenumber is derived by linearising the conservation of wave crests
(\ref{eq:wave_crests}) and neglecting the spatial gradients of $k$ and $g$.
The effective gravity is derived by projecting the long-wave induced surface
accelerations onto the local vertical axis (that is, the axis normal to the
long-wave surface), akin to \citet{zhang1990evolution}.
The amplitude is derived by first evaluating the short-wave action from the
action conservation (\ref{eq:wave_action}) and then applying the modulated
effective gravity to obtain the amplitude.
From the modulated wavenumber, amplitude, and effective gravity, the steepness,
frequency, and phase speed readily follow from the usual wave relationships.
Going forward we will use the tilde over the variables to denote the modulated
quantities (\textit{e.g.} $\widetilde{k}$ for wavenumber).

@@

\begin{equation}
\label{eq:wavenumber2}
\dfrac{\partial k}{\partial t}
= - k \dfrac{\partial U}{\partial x}
\end{equation}
Although \citet{peureux2021unsteady} pointed out that the solution by
\citet{longuet1960changes} requires assuming homogeneous group speed of the
short-wave field, in fact, it also requires assuming no horizontal propagation
and advection of the short waves.
Combine (\ref{eq:U_L})-(\ref{eq:wavenumber2}), and integrate in time to get:

@@

\begin{equation}
\label{eq:k_short_lhs}
\frac{\widetilde{k}}{k} = 1 + \varepsilon_L \cos{\psi}
\end{equation}
Coming from the conservation of crests, as opposed to the velocity potential
superposition, the solution by \citet{longuet1960changes} requires two
approximations:
First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
second, further truncating the Taylor expansion series beyond the first-order
term.
These approximations underestimate the short-wave modulation magnitude
(Fig. \ref{fig:analytical_solutions_ak0.1}).
At the wave crest, (\ref{eq:k_short_exact}) yields the wavenumber modulation
that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of
\citet{longuet1960changes} for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
Although not immediately obvious from the expressions, the modulated wavenumber
(\ref{eq:k_short_lhs}) is conserved across the long-wave phase, whereas
the higher-order solution (\ref{eq:k_short_exact}) is not.
This is because combining the orbital velocities evaluated at $z=\eta_L$ and the
linearised conservation of wave crests in (\ref{eq:wavenumber}) creates excess
short-wave wavenumber.
The solution of (\ref{eq:wavenumber}) must be conservative across the long-wave
phase, as we will see later from the numerical simulations.
The steady analytical solutions for modulation of short-wave effective gravity,
amplitude, steepness, intrinsic frequency, and phase speed are also shown in
Fig. \ref{fig:analytical_solutions_ak0.1}, for reference, however, their
formulations appear in the following subsections.
Although the analytical solutions by \citet{longuet1960changes} have been updated
by the numerical results of \citet{longuet1987propagation}, they are here
nevertheless relevant to use as reference for the proposed analytical solutions.
This comparison allows quantifying solely the effect of evaluating the
long-wave properties at the surface rather than at the mean water level.

@@

\begin{equation}
\label{eq:gravity_modulation_lagrangian}
\widetilde{g_l} =
g + \frac{dW}{dt} =
g + \frac{\partial W}{\partial t} + \left(C_g + U\right) \frac{\partial W}{\partial x} =
\widetilde{g_e} + \left(C_g + U\right) \frac{\partial W}{\partial x}
\end{equation}
where $l$ and $e$ subscripts serve to denote "Lagrangian" and "Eulerian",
respectively.
Although \citet{longuet1986eulerian} thoroughly described the differences
between the Eulerian and Lagrangian effective gravities, he did so for
a fully nonlinear irrotational gravity wave, which can be computed numerically
but not analytically.
Note also that when evaluating the Lagrangian effective gravity,
\citet{longuet1987propagation} considered the orbital motion of the long wave
but not the propagation speed of the short-wave group, which can exceed the
ambient velocity $U$ depending on the properties of the short and long waves
considered.
Group speed of the short waves aside, the Lagrangian correction to the gravity
being an advective term is $\mathcal{O}(\varepsilon_L^2)$.
Third, the projection of the long-wave induced horizontal acceleration
contributes small but non-negligible effects on the short-wave effective
gravity in the curvilinear coordinate system
\citep{phillips1981dispersion,zhang1990evolution}.
These three effects are completely described by projecting the gravitational
acceleration vector from the coordinate that is perpendicular to the mean water
surface ($z=0$) to that which is perpendicular to the long-wave surface
($z=\eta_L$), and likewise for the long-wave orbital accelerations:

@@

\begin{equation}
\label{eq:gravity_modulation_linear}
\frac{\widetilde{g_e}}{g}
  = 
  \frac{
    1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
    \left[ 1 + \left(\varepsilon_L \sin{\psi}\right)^2 e^{\varepsilon_L \cos{\psi}} \right]
  }
  {\sqrt{\left(\varepsilon_L \sin{\psi}\right)^2 + 1}}
\end{equation}
The same derivation can be carried out for the Lagrangian gravity
modulation following (\ref{eq:gravity_modulation_lagrangian})-(\ref{eq:local_slope}),
and is omitted here for brevity.
Curvilinear effects on effective gravity can be removed altogether by setting
$\alpha = 0$.
In that case, (\ref{eq:gravity_modulation_general}) simplifies to:

@@

The general form of the gravity modulation (\ref{eq:gravity_modulation_general})
is convenient because it allows us to easily attribute the modulation to
different processes.
It also makes it straightforward to evaluate the gravity modulation for a
different long-wave form, by evaluating the orbital accelerations and the local
slope for that wave.
For a third-order Stokes wave, for example, the elevation is:

@@

It is instructive to also examine the minimum values of the
effective gravity as function of long-wave steepness $\varepsilon_L$
(Fig. \ref{fig:effective_gravities_min}), as the minima occur at or
near the long-wave crests where the short-wave modulation is largest.
Without the Lagrangian correction to the effective gravity, the gravity
modulation is overestimated; for accelerations evaluated at the surface of a
linear wave, the maximum gravity reduction is 60\% at $\varepsilon_L = 0.4$.
The Lagrangian correction attenuates the gravity modulation, making the
gravity reduction at the crests $\approx 25\%$.
As the slope effects on the effective gravity exist only away from crests and
troughs, they may be neglected in the steady solutions where modulation is
typically evaluated at the long-wave crests.
Here, the slope effect causes a difference in the minimum effective gravity
for $\varepsilon_L \gtrsim 0.3$.
The effective gravity reduction at the crests in the case of a fully nonlinear
wave is $\approx$ 10\%, 19\%, 27\%, and 40\%, 
for $\varepsilon_L$ of 0.1, 0.2, 0.3, and 0.4, respectively.
These are similar to the fully nonlinear numerical estimates of
\citet{longuet1986eulerian,longuet1987propagation}.

@@

The homogeneity and stationarity of a scalar quantify $q$ can be expressed as
conditions on the instantaneous fractional rate of change of $q$ being much
smaller than the short-wave inverse scales:

@@

Now, to address the variation of short-wave properties,
we apply (\ref{eq:homogeneity_condition})-(\ref{eq:stationarity_condition}) to
the long-wave phase dependent wavenumber, gravitational acceleration, and
wave action, and define the homogeneity and stationarity quantities as:

>>>>

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that affects remote sensing, the
interpretation of \textit{in situ} wave measurements, and numerical wave
forecasting.
In this paper, we revisit the wave crest and action conservation laws and derive
higher-order steady solutions for the change of short-wave wavenumber, action,
and gravitational acceleration due to the presence of longer waves.
We validate the analytical solutions with numerical simulations of the full wave
crest and action conservation equations.
The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
steepness modulation significantly deviate from the linear analytical solutions
of \citet{longuet1960changes}, and are similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
The short-wave steepness modulation is attributed to be 5/8 due to the
wavenumber, 1/4 due to the wave action, and 1/8 due to the effective gravity.
Examining the homogeneity and stationarity requirements for the conservation of
wave action reveals that stationarity is a stronger requirement and is
generally not satisfied for very steep long waves.
We further examine the results of \citet{peureux2021unsteady} who found through
numerical simulations that the short-wave modulation grows unsteadily with
each long-wave passage.
We show that this unsteady growth only occurs for homogeneous initial
conditions as a special case and does not generally occur, for example, in more
realistic long-wave groups.
The proposed steady solutions are a good approximation of the nonlinear
crest-action conservation solutions in long-wave steepness $\lesssim 0.2$.
Except for a subset of initial conditions, the solutions to the non-linearised
crest-action conservation equations are mostly steady in the reference frame of
the long waves.
\end{abstract}

@@

\begin{equation}
\label{eq:lhs1960wavenumber}
\widetilde{k} = k (1 + \varepsilon_L \cos{\psi})
\end{equation}
where $\widetilde{k}$ and $k$ are the modulated and unmodulated short-wave wavenumbers,
respectively, $\varepsilon_L = a_L k_L$ is the steepness of the long wave with
wavenumber $k_L$ and amplitude $a_L$, and $\psi$ is the long-wave phase.
This process is illustrated in Fig. \ref{fig:hydrodynamic_modulation_diagram}.
Hydrodynamic modulation is a key process for the development and
interpretation of remote sensing of ocean surface waves and currents
\citep{keller1975microwave,alpers1978two,hara1994hydrodynamic} and is distinguished
from aerodynamic modulation \citep{donelan1987effect,belcher1999wave,chen2000effects}.
Experimental studies of the general short-wave modulation problem included both
the laboratory \citep[e.g.,][]{keller1975microwave,donelan2010modulation} and field
\citep[e.g.,][]{hara2003observation,plant1977studies} measurements.
\citet{phillips1981dispersion} and \citet{longuet1987propagation} revisited the
hydrodynamic modulation problem by considering nonlinear long waves and
variations in the effective gravity of short waves.
Their results yielded significantly stronger modulation than previously predicted
by \citet{longuet1960changes}.
\citet{henyey1988energy} derived an analytical solution for the modulation of
short waves using Hamiltonian mechanics and reported similar modulation magnitudes
to that of \citet{longuet1987propagation}.
\citet{zhang1990evolution} considered weakly nonlinear short waves
using a nonlinear Schr\" odinger equation and found stronger wavenumber
modulation but weaker amplitude modulation than those predicted by
\citet{longuet1987propagation}.
Aside from the early linear solutions by \citet{longuet1960changes}, the variety
of analytical and numerical frameworks (Eulerian crest and action conservation,
Hamiltonian, Schr\" odinger) that tackle the same fundamental
physics and that relatively closely reproduce the steady solutions for short-wave
modulation suggests that the solutions are robust.
Nevertheless, alternative analytical and numerical approaches to the modulation
problem remain of interest, especially if higher degrees of nonlinearity can be
considered using simpler approaches.

@@

All modulation solutions mentioned thus far are steady in the reference frame
of the long wave---short waves become shorter and higher (and thus, steeper)
preferentially on the long-wave crests.
Conversely, they elongate and flatten preferentially in the long-wave troughs.
Recently, \citet{peureux2021unsteady} asked whether the steady
(\textit{i.e.}, stationary; not depending on time) solutions
of short-wave modulation are appropriate by simulating the full wave conservation
equations.
They found that the short waves grow unsteadily due to the propagation of
longer waves, suggesting that the steady-state solutions may only be valid for
a few long-wave periods before the short waves begin to break due to excessive
steepness.
However, the unsteadiness of their solutions occurs in a specific scenario in
which a long-wave train suddenly appears to modulate a uniform field of short
waves.
The numerical simulations stabilise if the short-wave field is initialised
from existing steady solutions, like those by \citet{longuet1960changes}.
Nevertheless, their results do put into question whether the short-wave
modulation predicted by theory is generally steady over many long-wave periods
or not.
Looking at the conservation equations alone, it is not obvious that it is.

\begin{table}
\begin{center}
\def~{\hphantom{0}}
\begin{tabular}{cccccc}
Reference   & Analysis   & Solution & Solution    & Nonlinear  & Nonlinear   \\
            & framework  & type     & steadiness  & long waves & short waves \\
\hline
\citet{longuet1960changes} & Velocity potential  & Analytical & Steady & No  & No  \\
\hline
\citet{phillips1981dispersion} & Crest and action & Numerical & Steady & Yes & No  \\
                               & conservation     &           &        &     &     \\
\hline
\citet{longuet1987propagation} & Crest and action & Numerical & Steady & Yes & No  \\
                               & conservation     &           &        &     &     \\
\hline
\citet{henyey1988energy} & Hamiltonian & Analytical & Steady & Yes & No  \\
\hline
\citet{zhang1990evolution} & Schr\" odinger & Numerical & Steady & Yes & Yes \\
\hline
\citet{peureux2021unsteady} & Crest and action & Numerical & Unsteady & No  & No  \\
                            & conservation     &           &          &     &     \\
\hline
This paper  & Crest and action & Analytical    & Steady       & Yes & No  \\
            & conservation     & and numerical & and unsteady &     &     \\
\hline
\end{tabular}
\caption{
  Non-exhaustive summary of prior and present (this paper) approaches to
  calculating the modulation of short waves by long waves.
  The solution steadiness refers to whether the solution depends on time
  (unsteady) or not (steady).
}
\label{table:modulation_literature_summary}
\end{center}
\end{table}

In this paper, we revisit this problem from the perspective of wave action
conservation but consider the long-wave slope to be significant enough to
require evaluating the short-waves at the long-wave surface rather than at the
mean water level.
This allows us to derive alternative, steady, nonlinear solutions for the
modulation of short waves, within the limits of small $\varepsilon_L$.
The analytical solutions yield modulations that are stronger than the original
solution by \citet{longuet1960changes}, but similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution} for
$\varepsilon_L \lesssim 0.2$.
To validate the steadiness of the analytical solutions, we perform numerical
simulations of the full wave crest and action conservation laws, and compare
the results to the analytical solutions presented here, as well as the prior
numerical results.
The numerical simulations are used to determine the validity of the approximate
steady solution and to investigate the unsteadiness of the short-wave modulation
found by \citet{peureux2021unsteady}.
The main prior studies on hydrodynamic modulation, and their assumptions and
approaches to the solution, are summarised in Table
\ref{table:modulation_literature_summary}.
We first present the governing equations in \S\ref{section:governing_equations},
and then the analytical solutions for the modulation of short waves by long waves
in \S\ref{section:analytical_solutions}.
Then, we describe the numerical simulations and their results in
\S\ref{section:numerical_solutions}.
Finally, we discuss the results and their implications in
\S\ref{section:discussion} and conclude the paper in \S\ref{section:conclusions}.

@@

\begin{equation}
\label{eq:dispersion}
\omega = \sqrt{gk} + k U
\end{equation}
where $\omega$ is the angular frequency in a fixed reference frame, $g$ is the
gravitational acceleration, $k$ is the wavenumber, and $U$ is the mean advective
current in the direction of the wave propagation.
A current in the direction of the waves increases their absolute frequency
without changing their wavenumber.
An important limitation here is that $U$ must be slowly varying on the scales of
the wave period \citep{bretherton1968wavetrains}.
From the perspective of short waves riding on long waves, the advective current
is the horizontal near-surface orbital velocity of the long wave.
Strictly, a velocity profile must be considered when determining the
advective velocity for a wave of any given wavenumber \citep{stewart1974hf},
however, for brevity, here we will assume a surface velocity to be the advective
one for short waves.
The evolution of the short-wave wavenumber is described by the conservation of
wave crests \citep[e.g.,][]{whitham1974dispersive,phillips1981dispersion}:

@@

\begin{equation}
\label{eq:wave_action}
\dfrac{\partial N}{\partial t}
+ \dfrac{\partial}{\partial x} \left[\left(C_g + U\right)N\right]
= 0
\end{equation}
where $N$ is the action of short waves defined as the ratio of their
energy to their intrinsic frequency, $E/\sigma$,
and $C_g = 1/2\sqrt{g/k}$ is their group speed in deep water.
The wave action expressed as $E/\sigma$ is commonly taken as
conservative in wave modulation theory
\citep{phillips1981dispersion,longuet1987propagation}, however as
\citet{dysthe1988orbiting} show, it is conservative only for gently sloped waves.
We describe this and other requirements in more detail further below and show
in what long-wave conditions they are satisfied.
The wave crest conservation (\ref{eq:wave_crests}) follows from the assumption
of a conserved wave phase, which is limited by the same set of requirements as
those for the wave action conservation \citep[e.g.,][]{whitham1974dispersive}.
Although the absence of wave growth and dissipation is not generally realistic in
the open ocean, it is a useful approximation here to isolate the effects
of hydrodynamic modulation solely due to the motion of the long waves.

@@

\begin{equation}
\label{eq:wave_action2}
\dfrac{\partial N}{\partial t}
+ \left(C_g + U\right) \dfrac{\partial N}{\partial x}
+ N \dfrac{\partial U}{\partial x}
+ N \dfrac{\partial C_g}{\partial x}
= 0
\end{equation}
This equation is similar to (\ref{eq:wavenumber}), except for the last term
which represents the inhomogeneity of the group speed.
As \citet{peureux2021unsteady} explained, the presence of this term causes
unsteady growth of wave action in infinite long-wave trains.
(\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are the governing equations
that we approximate and solve analytically in \S\ref{section:analytical_solutions},
and numerically in their full form in \S\ref{section:numerical_solutions}.
This system of equations is semi-coupled, meaning that
it is possible to solve (\ref{eq:wavenumber}) on its own, but solving
(\ref{eq:wave_action2}) requires also the solution of (\ref{eq:wavenumber}).
In other words, the wave kinematics (the wavenumber distribution) are not
concerned with the wave action.
In contrast, the wave action is governed, among others, by the
group velocity and thus depends on the wavenumber.

@@

We now describe the analytical solutions for the short-wave wavenumber,
effective gravitational acceleration, amplitude, and steepness in the presence
of long waves.
The wavenumber is derived by linearising the conservation of wave crests
(\ref{eq:wave_crests}) and neglecting the spatial gradients of $k$ and $g$.
The effective gravity is derived by projecting the long-wave induced surface
accelerations onto the local vertical axis (that is, the axis normal to the
long-wave surface), akin to \citet{zhang1990evolution}.
The amplitude is derived by first evaluating the short-wave action from the
action conservation (\ref{eq:wave_action}) and then applying the modulated
effective gravity to obtain the amplitude.
From the modulated wavenumber, amplitude, and effective gravity, the steepness,
frequency, and phase speed readily follow from the usual wave relationships.
Going forward we will use the tilde over the variables to denote the modulated
quantities (\textit{e.g.} $\widetilde{k}$ for wavenumber).

@@

\begin{equation}
\label{eq:wavenumber2}
\dfrac{\partial k}{\partial t}
= - k \dfrac{\partial U}{\partial x}
\end{equation}
Although \citet{peureux2021unsteady} pointed out that the solution by
\citet{longuet1960changes} requires assuming homogeneous group speed of the
short-wave field, in fact, it also requires assuming no horizontal propagation
and advection of the short waves.
Combine (\ref{eq:U_L}) and (\ref{eq:wavenumber2}) and integrate in time to get:

@@

\begin{equation}
\label{eq:k_short_lhs}
\frac{\widetilde{k}}{k} = 1 + \varepsilon_L \cos{\psi}
\end{equation}
Coming from the conservation of crests, as opposed to the velocity potential
superposition, the solution by \citet{longuet1960changes} requires two
approximations:
First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
second, truncating the Taylor expansion series beyond $\mathcal{O}(\varepsilon_L)$.
These approximations underestimate the short-wave modulation magnitude
(Fig. \ref{fig:analytical_solutions_ak0.1}).
At the wave crest, (\ref{eq:k_short_exact}) yields the wavenumber modulation
that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of
\citet{longuet1960changes} for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
Although not immediately obvious from the expressions, the modulated wavenumber
(\ref{eq:k_short_lhs}) is conserved across the long-wave phase, whereas
the higher-order solution (\ref{eq:k_short_exact}) is not.
This is because combining the orbital velocities evaluated at $z=\eta_L$ and the
linearised conservation of wave crests in (\ref{eq:wavenumber}) creates excess
short-wave wavenumber.
The solution of (\ref{eq:wavenumber}) must be conservative across the long-wave
phase, as we will see later from the numerical simulations.
The steady analytical solutions for modulation of short-wave effective gravity,
amplitude, steepness, intrinsic frequency, and phase speed are also shown in
Fig. \ref{fig:analytical_solutions_ak0.1}, for reference, however, their
formulations appear in the following subsections.
Although the analytical solutions by \citet{longuet1960changes} have been updated
by the numerical results of \citet{longuet1987propagation}, they are here
nevertheless relevant to use as reference for the proposed analytical solutions.
This comparison allows quantifying solely the effect of evaluating the
long-wave properties at the surface rather than at the mean water level.

@@

\begin{equation}
\label{eq:gravity_modulation_lagrangian}
\widetilde{g_l} =
g + \frac{dW}{dt} =
g + \frac{\partial W}{\partial t} + \left(C_g + U\right) \frac{\partial W}{\partial x} =
\widetilde{g_e} + \left(C_g + U\right) \frac{\partial W}{\partial x}
\end{equation}
where $l$ and $e$ subscripts serve to denote "Lagrangian" and "Eulerian",
respectively.
Although \citet{longuet1986eulerian} thoroughly described the differences
between the Eulerian and Lagrangian effective gravities, he did so for
a fully-nonlinear irrotational gravity wave, which can be computed numerically
but not analytically.
Note also that when evaluating the Lagrangian effective gravity,
\citet{longuet1987propagation} considered the orbital motion of the long wave
but not the propagation speed of the short-wave group, which can exceed the
ambient velocity $U$ depending on the properties of the short and long waves
considered.
Group speed of the short waves aside, the Lagrangian correction to the gravity
being an advective term is $\mathcal{O}(\varepsilon_L^2)$.
Third, the projection of the long-wave induced horizontal acceleration in the
curvilinear coordinate system contributes small but non-negligible effects on
the short-wave effective gravity \citep{phillips1981dispersion,zhang1990evolution}.
These three effects are completely described by projecting the gravitational
acceleration vector from the coordinate that is perpendicular to the mean water
surface ($z=0$) to that which is perpendicular to the long-wave surface
($z=\eta_L$), and likewise for the long-wave orbital accelerations:

@@

\begin{equation}
\label{eq:gravity_modulation_linear}
\frac{\widetilde{g_e}}{g}
  = 
  \frac{
    1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
    \left[ 1 + \left(\varepsilon_L \sin{\psi}\right)^2 e^{\varepsilon_L \cos{\psi}} \right]
  }
  {\sqrt{\left(\varepsilon_L \sin{\psi}\right)^2 + 1}}
\end{equation}
The same derivation can be carried out for the Lagrangian gravity
modulation following (\ref{eq:gravity_modulation_lagrangian})-(\ref{eq:local_slope}),
and is omitted here for brevity.
Curvilinear effects on effective gravity can be removed altogether by setting
$\alpha = 0$.
In that case, (\ref{eq:gravity_modulation_linear}) simplifies to:

@@

The general form of the gravity modulation (\ref{eq:gravity_modulation_general})
is convenient because it allows us to easily attribute the modulation to
different processes.
It also makes it straightforward to evaluate the gravity modulation for
different long-wave forms, by evaluating the orbital accelerations and the local
slope for each wave form.
For a third-order Stokes wave, for example, the elevation is:

@@

It is instructive to also examine the minimum values of the
effective gravity as function of long-wave steepness $\varepsilon_L$
(Fig. \ref{fig:effective_gravities_min}), as the minima occur at or
near the long-wave crests where the short-wave modulation is largest.
Without the Lagrangian correction to the effective gravity, the gravity
modulation is overestimated; for accelerations evaluated at the surface of a
linear wave, the maximum gravity reduction is 60\% at $\varepsilon_L = 0.4$.
The Lagrangian correction attenuates the gravity modulation, making the
gravity reduction at the crests $\approx 25\%$.
As the slope effects on the effective gravity exist only away from crests and
troughs, they may be neglected in the steady solutions where modulation is
typically evaluated at the long-wave crests.
Here, the slope effect causes a difference in the minimum effective gravity
for $\varepsilon_L \gtrsim 0.3$.
The effective gravity reduction at the crests in the case of a fully-nonlinear
wave is $\approx$ 10\%, 19\%, 27\%, and 40\%, 
for $\varepsilon_L$ of 0.1, 0.2, 0.3, and 0.4, respectively.
These are similar to the fully-nonlinear numerical estimates of
\citet{longuet1986eulerian,longuet1987propagation}.

@@

The homogeneity and stationarity of a scalar quantity $q$ can be expressed as
conditions on the instantaneous fractional rate of change of $q$ being much
smaller than the short-wave inverse scales:

@@

Now, to address the variation of short-wave properties,
we apply (\ref{eq:homogeneity_condition})-(\ref{eq:stationarity_condition}) to
the long-wave phase dependent wavenumber, gravitational acceleration, and
wave action, and define the homogeneity and stationarity as:
